Here is another notebook for data analyses

```{r}
library(MuMIn)
library(readr)
library(corrplot)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyverse)
library(psych)
library(readxl)
library(lmerTest)
library(MuMIn)
library(factoextra)
library(simr)
library(car)
library(sjPlot)
```

Demographics (mean age, sd age)

```{r}
demo_data_clean %>%
  summarize(mean(age),
            sd(age))
```



MM1 PREP DF
```{r}

# PREPARE A DF FOR MM1 Assuming your data frame is named mm1prep
mm1prepping <- ratings %>%
  select(exp_subject_id, audio_name, monra) %>%
  pivot_wider(names_from = audio_name, values_from = monra)

```


Calculate MM1 Values
```{r}

MM1_revised <- function(df){
  #save & remove subj id
  id <- df[, 1]
  df <- as.matrix(df[,-1])
  
  #Compute means minus one
  means <- matrix(0, nrow = nrow(df), ncol = ncol(df))
  for(j in 1:nrow(df)){
    means[j,] <- colMeans(df[-j,])
  }
  
  #Compute correlations between means minus ones and ones ratings
  mm1 <- sapply(1:nrow(df), function(.x) cor(df[.x,], means[.x,]))
  
  #transform correlations into z scores (r to z Fisher)
  z_scores <- sapply(mm1, function(x) log((1 + x)/(1 - x))) * .5 
  
  #save average correlation (z average to r average Fisher) and correlations per rater
  result <-  list(MM1 = (exp(2 * (mean(z_scores))) - 1)/(exp( 2 * (mean(z_scores))) + 1), 
                  raw = data.frame(id, mm1 = mm1, mm1_z = z_scores))
  result
}

#I wanted to see the MM1 list pls
MM1_List <- MM1_revised(mm1prepping)
MM1_L <- as.data.frame(MM1_List$raw)
MM1_List
MM1_List$MM1
```


What is the mean of all monra ratings across 9000 trials
```{r}
monra_data %>%
  summarize(mean(monra))
```

What is the mean rating of stimuli that are rated as music
```{r}
# Calculate the mean of 'monra' for 'PAG' M
mean_monra_M <- mean(ratings$monra[ratings$PAG == 'M'], na.rm = TRUE)

# Calculate the mean of 'monra' for 'PAG' A
mean_monra_A <- mean(ratings$monra[ratings$PAG == 'A'], na.rm = TRUE)

# Calculate the mean of 'monra' for 'PAG' N
mean_monra_N <- mean(ratings$monra[ratings$PAG == 'N'], na.rm = TRUE)

print(mean_monra_M)
print(mean_monra_A)
print(mean_monra_N)
```


Now let's do a Linear Mixed Model,  by using PAG as fixed effects and audio names and participants as random
```{r}
# Fit a mixed-effects model
model <- lmer(monra ~ PAG + (1|exp_subject_id) + (1 | audio_name), data = ratings)

# View the summary of the model
summary(model)
anova_results <- anova(model)
r.squaredGLMM(model)
```



We'll try to calculate what is the best number of clusters for this dataset,
We do this by using the Elbow Method

```{r}
# Look at clusters with a defined number of clusters
kmm = kmeans(audio_descriptives$mean,3,nstart = 50,iter.max = 15) #we keep number of iter.max=15 to ensure the algorithm converges and nstart=50 to #ensure that atleat 50 random sets are choosen  
kmm

## Methods 1: Elbow Method for finding the optimal number of clusters
set.seed(123)
# Compute and plot wss for k = 2 to k = 15.
k.max <- 15
data_for_kmm <- audio_descriptives$mean
wss <- sapply(1:k.max, 
              function(k){kmeans(data_for_kmm, k, nstart=50,iter.max = 15 )$tot.withinss})
wss
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")

```
As observed here, 3 seems to be a valid option here.

```{r}

slider_long <- ratings %>%
  select(exp_subject_id, audio_name, monra) %>%
  mutate(condition = "slider",
         monra = ifelse(monra >= 50, 1, 0))

##this is the data from experiment 1

fcself <- read_excel("preprocessed_data_all.xlsx", 
    sheet = "Figures_self")

afc_slider <- fcself %>%
  rename(exp_subject_id = Participant,
         monra = Music_answer,
         audio_name = Stim) %>%
  select(exp_subject_id, 
         audio_name, monra) %>%
  mutate(condition = "afc") %>%
  bind_rows(slider_long)

afc_slider$monra <- as.numeric(afc_slider$monra)
afc_slider$condition <- as.factor(afc_slider$condition)
afc_slider$condition <- factor(afc_slider$condition, levels = c("slider", "afc"))
str(afc_slider)

library(emmeans)

# Fit a mixed-effects model
model_comp <- glmer(monra ~ condition + (1 |exp_subject_id) + (1 | audio_name), data = afc_slider, family = "binomial")

summary(model_comp)
r.squaredGLMM(model_comp)
performance::icc(model_comp, by_group = TRUE)



```

Correlation between the two conditions

```{r}
afc_slider

audio_descriptives_scaled$mean <- as.numeric(audio_descriptives_scaled$mean)
PAGG$Music_answer <- as.numeric(PAGG$Music_answer)

cor.test(audio_descriptives_scaled$mean, PAGG$Music_answer)

```

