R script to exclude participants who failed the attention check, or did not complete the study

```{r}

library(dplyr)
library(tibble)
library(writexl)
library(janitor)
library(readxl)
```

Load the data

```{r}

untouched_data <- read_xlsx("all_data_new.xlsx",
                  sheet = "all_data_new")

data <- read_xlsx("all_data_new.xlsx",
                  sheet = "all_data_new")

data <- janitor::clean_names(data)

data <- data %>%
  rename(monra = musicornot)

data %>%
  distinct(exp_subject_id) #120

data %>%
  distinct(prolific_id) #113 + (na)
```


2 PAIRS (797627,797719 AND 797633,797684) HAVE THE SAME PROLIFIC ID,THAT'S WHERE THE INCONSISTENCY COMES FROM 

```{r}
data %>%
  filter(task_name == "Instructions") %>%
  distinct(exp_subject_id, prolific_id)

```



Come up with audio groups as categorization

```{r}

data <- data %>%

  mutate(categorization = case_when(

    startsWith(audio_name, "M_") ~ "music",

    startsWith(audio_name, "N_") ~ "not_music",

    startsWith(audio_name, "SIGNAL_") ~ "signal",

    TRUE ~ NA_character_

  ))



```

 
```{r}

prolific_id <- data %>%
  filter(task_name == "Instructions") %>%
  select(exp_subject_id, prolific_id)

demo_data_everyone <- data %>% 
  filter(task_name == "Demographic_data") %>% 
  dplyr::select(exp_subject_id, age, gender, hearing_impairments,
         browser_spec, completed, crowdsourcing_code, crowdsourcing_subj_id, start_time, start_time_local, end_time, end_time_local, 
         session_nr, session_token, group_nr) %>%
  left_join(prolific_id %>%
              select(exp_subject_id, prolific_id),
            by = "exp_subject_id")

demo_data_clean <- demo_data_everyone


```

Filtering out participants who didn't complete, who had hearing impairments, who didn't consent
 
```{r}

incomplete <- data %>%
  filter(completed != TRUE) %>% 
  dplyr::select(exp_subject_id) %>% 
  distinct(exp_subject_id) %>%
  deframe()
incomplete

# Now drop incomplete datasets

data <- data %>% filter(!exp_subject_id %in% incomplete)
demo_data_clean <- demo_data_clean %>% filter(!exp_subject_id %in% incomplete)

# Exclude participants that did not consent
no_consent <- data %>% 
  filter(task_name == "Informed_consent" & informed_consent != TRUE) %>% 
  dplyr::select(exp_subject_id) %>% 
  distinct(exp_subject_id) %>%
  deframe()
no_consent


data <- data %>% filter(!exp_subject_id %in% no_consent)
demo_data_clean <- demo_data_clean %>% filter(!exp_subject_id %in% no_consent)


# Exclude participants that have hearing impairments
hearing_impairement <- data %>%
  filter(task_name == "Demographic_data" & hearing_impairments != "No") %>% 
  dplyr::select(exp_subject_id) %>% 
  distinct(exp_subject_id) %>%
  deframe()
hearing_impairement


data <- data %>% filter(!exp_subject_id %in% hearing_impairement)
demo_data_clean <- demo_data_clean %>% filter(!exp_subject_id %in% hearing_impairement)



```
 

Compute the mean monra for all the participants, make it appear as a seperate column. Come up with bark_data 2 where you have 1 row for every participant and columns for their participant ID, mean monra and their completion status. PROLIFIC ID WILL BE ADDED

```{r}


bark_data <- filter(data, categorization %in% c("signal"))

bark_data2 <- bark_data %>%
  group_by(exp_subject_id) %>%
   summarise(
    mean_monra = mean(monra, na.rm = TRUE))

demo_data_everyone <- merge(bark_data2, demo_data_everyone, by = "exp_subject_id")
demo_data_clean <- merge(bark_data2, demo_data_clean, by = "exp_subject_id")


no_attention <- bark_data2 %>% 
  group_by(exp_subject_id) %>% 
  filter(mean_monra < 80) %>% 
  dplyr::select(exp_subject_id) %>% 
  deframe()
no_attention

data <- data %>% filter(!exp_subject_id %in% no_attention)
demo_data_clean <- demo_data_clean %>% filter(!exp_subject_id %in% no_attention)

no_attention
hearing_impairement
incomplete
no_consent

write_xlsx(data, "clean_data.xlsx")


data %>%
  filter(task_name == "Instructions") %>%
  distinct(exp_subject_id, prolific_id) ##95



```



