---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(ggplot2)
library(reshape2)
library(corrplot)

```

Plots for showing the correlations between BAIS-V and Gold-MSI and music ratings, not reported

```{r}
options(scipen = 999)  # Prevents scientific notation


correlation_value <- cor(scales$mean_rating, scales$baiss, use = "complete.obs")
print(correlation_value)

correlation_value2 <- cor(scales$mean_rating, scales$goldd, use = "complete.obs")
print(correlation_value2)

# Creating the plot
plot(scales$mean_rating, scales$baiss, pch = 19, col = "lightblue",
     xlab = "Mean Rating", ylab = "BAIS-V Score")

# Regression line
abline(lm(scales$baiss ~ scales$mean_rating), col = "red", lwd = 3)

# Pearson correlation
cor_value <- round(cor(scales$mean_rating, scales$baiss), 2)
text(x = min(scales$mean_rating), y = max(scales$baiss), 
     labels = paste("Correlation:", cor_value), pos = 4)


# Creating the plot
plot(scales$mean_rating, scales$goldd, pch = 19, col = "lightblue",
     xlab = "Mean Rating", ylab = "Gold-MSI Score")

# Regression line
abline(lm(scales$goldd ~ scales$mean_rating), col = "red", lwd = 3)

# Pearson correlation
cor_value <- round(cor(scales$mean_rating, scales$goldd), 2)
text(x = min(scales$mean_rating), y = max(scales$goldd), 
     labels = paste("Correlation:", cor_value), pos = 4)

```

Gradient plot, the first column represent the ranking of mean ratings at the stimulus level. This shows that even though participants at the group level, there is still some individual variability (reported in SI)

```{r}

# Reshaping the dataframe
monra_mm1_long <- mm1prepping %>%
  pivot_longer(
    cols = -exp_subject_id,  # Keep exp_subject_id, gather other columns
    names_to = "audio_name",  # The names of the audio columns will go to this new column
    values_to = "rating"      # The ratings will go to this new column
  )

# Now pivot wider so that exp_subject_id becomes columns
monra_mm1_wide <- monra_mm1_long %>%
  pivot_wider(
    names_from = exp_subject_id,  # The exp_subject_id values become new columns
    values_from = rating          # Values come from the rating column
  )

# Print the reshaped dataframe
print(monra_mm1_wide)

monra_mean_participants <- monra_mm1_wide %>%
  left_join(audio_descriptives %>% dplyr::select(audio_name, mean), by = "audio_name") %>%
  relocate(mean, .after = audio_name)

# Step 1: Reshape the original data to long format for participant ratings
long_data <- monra_mean_participants %>%
  pivot_longer(
    cols = -c(audio_name, mean),  # Exclude non-participant columns
    names_to = "exp_subject_id", 
    values_to = "rating"
  ) %>%
  group_by(exp_subject_id) %>%
  arrange(rating, .by_group = TRUE) %>%
  mutate(rank = row_number())  # Assign rank based on the order of ratings within each participant

# Step 2: Create a long format for mean_monra
mean_monra_long <- monra_mean_participants %>%
  dplyr::select(audio_name, mean) %>%
  mutate(exp_subject_id = "mean", rating = mean) %>%
  dplyr::select(audio_name, exp_subject_id, rating)

# Step 3: Combine both long formats
combined_long_data <- bind_rows(long_data, mean_monra_long)

# Step 4: Adjust the participant_id to make mean_monra first
combined_long_data <- combined_long_data %>%
  mutate(exp_subject_id = factor(exp_subject_id, levels = c("mean", unique(long_data$exp_subject_id))))  # Ensures mean_monra is first

# Step 5: Rank the combined data
combined_long_data <- combined_long_data %>%
  group_by(exp_subject_id) %>%
  arrange(rating, .by_group = TRUE) %>%
  mutate(rank = row_number())  # Assign rank for all participants including mean_monra
  
# Ensure the data has a gradient variable within each Audio_Name
combined_long_data <- combined_long_data %>%
  group_by(audio_name) %>%
  mutate(gradient_value = scales::rescale(rank))  # Rescale rank for gradient (0 to 1)

# Plot with gradients
ggplot(combined_long_data, aes(x = exp_subject_id, y = rank, fill = gradient_value)) +
  geom_tile() +
  scale_fill_gradientn(
    colors = c("beige", "red"),   # Gradient for music (modify for other categories)
    guide = "none"              # Hide legend for the gradient
  ) +
  labs(
    x = "Participants",
    y = "Ranking of Stimuli",
    title = " ",
    fill = "Audio_Name"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),  # Hide x-axis text labels
    axis.title.x = element_text(size = 20),
    axis.text.y = element_blank(),  # Hide y-axis text labels
    axis.title.y = element_text(size = 20),
    panel.grid = element_blank(),  # Remove grid lines
    legend.position = "none"       # Remove legend
  )

```


Correlation matrix of the group (i.e., forced-choice binary versus slider) and individual rankings (n = 95, slider only) of the 90 stimuli under study, reported

```{r}

mm1prepping_2 <- mm1prepping

mm1prepping_3 <- monra_mm1_long %>%
  pivot_wider(names_from = exp_subject_id, values_from = rating) %>%
  mutate(New_Names = str_replace(audio_name, '^[MN]_(.+)\\.mp3$', '\\1')) %>%
  mutate(New_Names = str_to_title(New_Names)) %>%
  select(New_Names, everything()) %>%
  select(-audio_name) %>%
  left_join(
    audio_descriptives_scaled %>%
      select(New_Names, Music_answer, mean),
    join_by(New_Names)
  ) %>%
  select(New_Names, Music_answer, mean, everything())

mm1prepping_4 <- monra_mm1_long %>%
  pivot_wider(names_from = exp_subject_id, values_from = rating) %>%
  mutate(New_Names = str_replace(audio_name, '^[MN]_(.+)\\.mp3$', '\\1')) %>%
  mutate(New_Names = str_to_title(New_Names)) %>%
  select(New_Names, everything()) %>%
  select(-audio_name) %>%
  left_join(
    audio_descriptives %>%
      select(New_Names, Music_answer, mean),
    by = "New_Names") %>%
  select(New_Names, Music_answer, mean, everything()) %>%
  rename(Slider = mean,
         Baseline = Music_answer)


# Select Music_answer, mean_rating, and all columns that have numeric names (participant ratings)
selected_columns <- mm1prepping_4 %>%
  select(Baseline, Slider, matches("^\\d+$"))

# Compute the correlation matrix
correlation_matrix <- cor(selected_columns, use = "complete.obs", method = "pearson")

# View the correlation matrix
print(correlation_matrix)

# Reshape the correlation matrix to long format
melted_correlation_matrix <- melt(correlation_matrix)

# Plot the heatmap using ggplot2
ggplot(data = melted_correlation_matrix, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "white", high = "red", mid = "green", 
                       midpoint = 0.5, limit = c(0, 1), space = "Lab", 
                       name="Correlation") +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(), # Hide x-axis title
    axis.title.y = element_blank(), # Hide y-axis title
    axis.ticks = element_blank(),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  # Rotate x-axis text to vertical
    axis.text.y = element_text(angle = 0, vjust = 0.5)# Hide axis ticks
  ) +
  coord_fixed() +
  labs(title = " ")



ggplot(data = melted_correlation_matrix, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(
    low = "white", high = "red", mid = "green", 
    midpoint = 0.5, limit = c(0, 1), space = "Lab", 
    name = "Correlation"
  ) +
  scale_x_discrete(labels = function(x) ifelse(x %in% c("Baseline", "Slider"), x, "")) +  # Show only "baseline" and "slider"
  scale_y_discrete(labels = function(y) ifelse(y %in% c("Baseline", "Slider"), y, "")) +  # Show only "baseline" and "slider"
  theme_minimal() +
  theme(
    axis.title.x = element_blank(), 
    axis.title.y = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(vjust = 0.5)
  ) +
  coord_fixed() +
  labs(title = " ")
```

Plot for showing mean music ratings in two studies, reported

```{r}

audio_descriptives_scaled$mean_rating <- audio_descriptives_scaled$mean
  
ggplot(audio_descriptives_scaled, aes(x = New_Names, y = mean_rating)) +
  geom_line(aes(x = reorder(New_Names, Music_answer), y = Music_answer, linetype = "Forced-choice"), group = 1) +  # Add linetype mapping for lines
  geom_point(size = 2, aes(shape = "Slider")) +  
  geom_errorbar(aes(ymin = mean - ci, ymax = mean + ci), width = 0.3) +
  labs(x = "Sound Excerpts", y = "Music Answers") +
  scale_shape_manual(name = " ", values = c("Slider" = 16), labels = c("Slider")) +  # Customize shape legend
  scale_linetype_manual(name = " ", values = c("Forced-choice" = "solid"), labels = c("Forced-choice")) +  # Customize linetype legend
  theme_minimal() +
  ylim(0, 1) +
  theme(axis.text.x = element_text(color = "black", angle = 90, vjust = 0.5, size = 10),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(color = "black", size = 30),
        axis.title.y = element_text(color = "black", angle = 90, size = 30),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "top")

cor.test(mm1prepping_4$Baseline, mm1prepping_4$Slider, method = "spearman")

```

Raw ratings plotted, with a geom smooth curve, colored by the clusters identified in experiment 1, not reported.

```{r}

ggplot(ratings, aes(x = reorder(audio_name, monra), y = monra, color = as.factor(PAG), group = 1)) +
  geom_jitter(alpha = 0.4) +
  geom_smooth(color = "black") +
  labs(title = " ",
       x = "Audios",
       y = "Music or Not Rating",
       color = "Predetermined Audio Groups") +
    scale_color_manual(values = c("A" = "blue", "M" = "red", "N" = "green")) +
  theme_minimal() +
  theme(axis.text.x = element_text(color = "black", angle = 90, vjust = 0.5, size = 9),
        axis.title.x = element_text(color = "black", size = 35),
        axis.title.y = element_text(color = "black", angle = 90, size = 35),
        axis.text.y = element_text(size = 15),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "none") 

```

