
```{r}
#install.packages("maditr")
library(maditr)
library(stringr)
library(hablar)
library(readxl)
library(dplyr)
library(writexl)

```

First I come up with the df called features_data, and have a more structured and clean table there
```{r}

ratings_scales <- dplyr::select(similarity_sep, exp_subject_id, task_name, audio_name, audio_group, similarity, ae_01, ae_02, ae_05, ae_07, bais_v_01, bais_v_02, bais_v_03, bais_v_04, bais_v_05, bais_v_06, bais_v_07, bais_v_08, bais_v_09, bais_v_10, bais_v_11, bais_v_12, bais_v_13, bais_v_14, mt_01, mt_02, mt_03, mt_06, mt_07, pa_04, pa_08, sa_01, sa_02, sa_03, sa_04, sa_05, sa_06, em_04, open1, open2, open3, open4, open5, open6, open7, open8, open9, open10)

```


Data Frames for MM1 analysis

```{r}

similarity_mm1 <- similarity_sep %>%
  select(exp_subject_id, audio_name, similarity) %>%
  pivot_wider(names_from = audio_name, values_from = similarity)

```



A dataframe for the means of all audios
```{r}

similarity_desc <- similarity_sep %>%
  group_by(audio_name) %>%
  summarize(
            mean_similarity = mean(similarity, na.rm = TRUE),
            se = sd(similarity, na.rm = TRUE) / sqrt(n()))

##with blocknames
similarity_desc2 <- similarity_desc %>%
  left_join(
    data %>%
      distinct(audio_name, block_name),  # Retain unique combinations
    by = "audio_name"
  )
library(writexl)
write_xlsx(similarity_desc2, "similarity_desc2.xlsx")

mean(similarity_clean$similarity)
sd(similarity_clean$similarity)


```




Calculate Gold-MSI
```{r}

gold <- dplyr::select(data, exp_subject_id, ae_01, ae_02, ae_05, ae_07, mt_01, mt_02, mt_03, mt_06, mt_07, pa_04, pa_08, sa_01, sa_02, sa_03, sa_04, sa_05, sa_06, em_04)

gold <- aggregated_data <- aggregate(. ~ exp_subject_id, data = gold, FUN = function(x) x[!is.na(x)][1])

gold2<-dplyr::select(gold, ! exp_subject_id)

keys_gold <- c("sa_05", "sa_04", "mt_03", "mt_07", "pa_08")

rec <-  psych::reverse.code(keys = keys_gold, items = gold2, mini = 1, maxi = 7) 

# View(rec)
rec <- as.data.frame(rec)

#GM_sum 
rec$gold=rowSums(cbind(rec$ae_01,rec$ae_02,rec$ae_05,rec$ae_07,rec$em_04,rec$mt_01,rec$mt_02,rec$`mt_03-`,rec$mt_06,rec$`mt_07-`,rec$pa_04,rec$`pa_08-`,rec$sa_01,rec$sa_02,rec$sa_03,rec$`sa_04-`,rec$`sa_05-`,rec$sa_06))

rec$exp_subject_id <- gold$exp_subject_id
gold_score <- dplyr::select(rec, exp_subject_id, gold)


```


Calculate BAIS-V
```{r}

bais <- dplyr::select(data, exp_subject_id, bais_v_01, bais_v_02, bais_v_03, bais_v_04, bais_v_05, bais_v_06, bais_v_07, bais_v_08, bais_v_09, bais_v_10, bais_v_11, bais_v_12, bais_v_13, bais_v_14)
bais <- na.omit(bais)
rec <- as.data.frame(bais)

#Mean_score
rec$bais=rowMeans(cbind(rec$bais_v_01,rec$bais_v_02,rec$bais_v_03,rec$bais_v_04,rec$bais_v_05,rec$bais_v_06,rec$bais_v_07,rec$bais_v_08,rec$bais_v_09,rec$bais_v_10,rec$bais_v_11,rec$bais_v_12,rec$bais_v_13,rec$bais_v_14))
rec$exp_subject_id <- bais$exp_subject_id

BAIS_score<-dplyr::select(rec,exp_subject_id, bais)

scales <-merge(gold_score, BAIS_score, by = "exp_subject_id")

```


Calculate Openness

```{r}

openness <- dplyr::select(data, exp_subject_id, open1, open2, open3, open4, open5, open6, open7, open8, open9, open10)

reverse_code <- function(x) {
  return(6 - x)
}

# Apply reverse coding and calculate total_openness
openness <- openness %>%
  mutate(
    open2 = reverse_code(open2),
    open4 = reverse_code(open4),
    open6 = reverse_code(open6),
    open8 = reverse_code(open8),
    open10 = reverse_code(open10)
  ) %>%
  group_by(exp_subject_id) %>%
  summarise(total_openness = sum(c_across(open1:open10), na.rm = TRUE))

# View the result
print(openness)

scales <- merge(scales, openness, by ="exp_subject_id")

```

Let's compute genres 

```{r}

genres <- data %>%
  dplyr::select(exp_subject_id, top40, pop, rock, classical, noise, experimental, edm) %>%
  na.omit() 

```


Calculate the mean age
```{r}

mean(demo_data_everyone$age)
sd(demo_data_everyone$age)


```


Data cleaning

Reading in acoustic information and coming up with similarity and dissimilarity matrices for future plotting

```{r}

acoustic_pc <- read_xlsx("acoustic_pc.xlsx",
                               sheet = "Sheet1")

acoustic_features <- read_xlsx("selected_acoustic.xlsx",
                               sheet = "Sheet1")
  
similarity_clean_sorted <- similarity_clean %>%
  mutate(
    audio1_sorted = pmin(audio1, audio2),
    audio2_sorted = pmax(audio1, audio2)
  )
similarity_matrix_sorted <- dcast(similarity_clean_sorted, audio2_sorted ~ audio1_sorted, value.var = "similarity", fun.aggregate = mean, fill = 0)
rownames(similarity_matrix_sorted) <- similarity_matrix_sorted$audio2_sorted
similarity_matrix_sorted$audio2_sorted <- NULL

similarity_matrix_sorted_matrix <- as.matrix(similarity_matrix_sorted)

###Create dfs to be able to do this. Since participants rated Stimulus1-Stimulus2 but not S2-S1, distance matrix is impossible to make. This is because how one half of the matrix is identical ot the other half... So we change the order of columns, duplicate it, and the bind it to the original df

similarity_clean1 <- similarity_clean %>%
  select(audio1, audio2, similarity)

similarity_clean2 <- similarity_clean %>%
  mutate(audio3 = audio1,
         audio4 = audio2) %>%
  select(audio4, audio3, similarity)

similarity_clean2 <- similarity_clean2 %>%
  rename(audio1 = audio4,
         audio2 = audio3)

similarity_double <- rbind(similarity_clean1, similarity_clean2)

dissimilarity_double <- similarity_double %>%
  mutate(distance = 100 - similarity) %>%
  select(audio1, audio2, distance)

dissimilarity <- similarity_double %>%
  mutate(distance = 100 - similarity) %>%
   mutate(
    audio1_sorted = pmin(audio1, audio2),
    audio2_sorted = pmax(audio1, audio2)
  )
distance_matrix <- dcast(dissimilarity_double, audio2 ~ audio1, value.var = "distance", fun.aggregate = mean, fill = 0)

similarity_matrix <- dcast(similarity_double, audio2 ~ audio1, value.var = "similarity", fun.aggregate = mean, fill = 0)


rownames(similarity_matrix) <- similarity_matrix$audio2

rownames(distance_matrix) <- distance_matrix$audio2

# Remove the first column from the matrix
distance_matrix <- distance_matrix[, -1]
similarity_matrix <- similarity_matrix[, -1]

distance_matrix <- as.matrix(distance_matrix)
similarity_matrix <- as.matrix(similarity_matrix)


# Assign row and column names to the similarity matrix
#rownames(similarity_matrix_sorted_matrix) <- stimuli_names
#colnames(similarity_matrix_sorted_matrix) <- stimuli_names


# Plot the heatmap without dendrograms and ordering

distance_matrix_flipped_horizontally <- distance_matrix[, ncol(distance_matrix):1]


distance_matrix2 <- distance_matrix

distance_matrix2[lower.tri(distance_matrix2)] <- NA

distance_matrix3 <- distance_matrix
distance_matrix3[upper.tri(distance_matrix3)] <- NA

similarity_matrix2 <- similarity_matrix
similarity_matrix2[upper.tri(similarity_matrix2)] <- NA

```


Let's calculate the mean similarity score per each audio

```{r}

mean_similarity1 <- similarity_clean %>%
   group_by(audio1) %>%
  summarize(mean_similarity1 = mean(similarity)) 
mean_similarity1$audio_name <- mean_similarity1$audio1

mean_similarity2 <- similarity_clean %>%
    group_by(audio2) %>%
  summarize(mean_similarity2 = mean(similarity)) 
mean_similarity2$audio_name <- mean_similarity2$audio2

mean_similarity <- merge(mean_similarity1, mean_similarity2, by = "audio_name") 

mean_similarity <- mean_similarity %>%
  group_by(audio_name) %>%
  mutate(mean_similarity = (mean_similarity1 + mean_similarity2) / 2) %>%
  select(audio_name, mean_similarity)
  



```


