
```{r}

library(tidyverse)
library(stringr)
library(hablar)
library(readxl)
library(dplyr)
library(writexl)

#install.packages("writexl")
#install.packages("hablar")

```

```{r}

setwd("C:/Users/talip/Desktop/MPIAE/Perceptual features study/data/similarity")

audio_features <- read_xlsx("audio_features.xlsx", sheet = "Sheet1")
audio_features <- audio_features %>%
  filter(label == "music") %>%
  select(!label)

data <- read_csv("trials_and_sessions.csv")

untouched_data <- data


ratings_lme <- read_xlsx("ratings_lme.xlsx", sheet = "Sheet 1")


```


Read in data
```{r}
data <- janitor::clean_names(data)

data %>% distinct(crowdsourcing_subj_id) 
data %>% distinct(exp_subject_id) 
data %>% distinct(prolific_id)     


participants <- data %>%
  filter(task_name == "Instructions") %>%
  distinct(exp_subject_id, prolific_id)


```


Make column names with lowercase letters and change column data types to logical

```{r}


data <- data %>% 
  hablar::convert(lgl(completed, 
              informed_consent,
              ends_with("_fam")
              )
          )




data <- data %>%

  mutate(audio_group = case_when(

    startsWith(audio_name, "M_") ~ "music",

    startsWith(audio_name, "N_") ~ "not_music",

    startsWith(audio_name, "SIGNAL_") ~ "signal",
  ))

```

Creating a DF for demographics of participants (IT DOESN'T SHOW 1 PARTICIPANT HERE BECAUSE OF INCOMPLETE DATA, SO IF U SEE 15 PARTICIPANT IN DEMO_DATA_EVERYONE, IT'S ACTUALLY 16)

```{r}

demo_data_everyone <- data %>% 
  filter(task_name == "Demographic_data") %>% 
  dplyr::select(exp_subject_id, age, gender, hearing_impairments, block_name,
         browser_spec, completed, crowdsourcing_code, crowdsourcing_subj_id, start_time, start_time_local, end_time, end_time_local, 
         session_nr, session_token, group_nr)

# join demo_data with columns from introduction, consent and additional questions
demo_data_everyone <- demo_data_everyone %>% 
  left_join(
    data %>% 
      filter(task_name == "Informed_consent") %>% 
      dplyr::select(exp_subject_id, informed_consent),
    join_by(exp_subject_id)
  ) %>% 
  left_join(
    data %>% 
      filter(task_name == "Honest_answers") %>% 
      dplyr::select(exp_subject_id, conscientious, wear_headphones),
    join_by(exp_subject_id)
  ) %>%
  left_join(
    data %>%
      filter(task_name == "Instructions") %>%
      dplyr::select(exp_subject_id, prolific_id),
    join_by(exp_subject_id)
  )



data %>%
  distinct(exp_subject_id)

data %>%
  distinct(prolific_id)

part_per_condition <- demo_data_everyone %>%
  group_by(block_name) %>%
  summarize(count = n())

```

######################
 EXCLUSION CRITERIA 
######################

```{r}

incomplete <- data %>%
  filter(completed != TRUE) %>% 
  dplyr::select(exp_subject_id) %>% 
  distinct(exp_subject_id) %>%
  deframe()
incomplete

data <- data %>% filter(!exp_subject_id %in% incomplete)

# Exclude participants that did not consent
no_consent <- data %>% 
  filter(task_name == "Informed_consent" & informed_consent != TRUE) %>% 
  dplyr::select(exp_subject_id) %>% 
  distinct(exp_subject_id) %>%
  deframe()
no_consent


data <- data %>% filter(!exp_subject_id %in% no_consent)


# Exclude participants that have hearing impairments?
hearing_impairement <- data %>%
  filter(task_name == "Demographic_data" & hearing_impairments != "No") %>% 
  dplyr::select(exp_subject_id) %>% 
  distinct(exp_subject_id) %>%
  deframe()
hearing_impairement


data <- data %>% filter(!exp_subject_id %in% hearing_impairement)




# Exclude participants who missed too many attention checks >80%

bark_data <- data %>%
  filter(audio_group == "signal") %>%
  group_by(exp_subject_id) %>%
  summarise(mean_similarity = mean(similarity, na.rm = TRUE)) %>%
  dplyr::select(exp_subject_id, mean_similarity) 

no_attention <- bark_data %>% 
  group_by(exp_subject_id) %>% 
  filter(mean_similarity < 80) %>% 
  dplyr::select(exp_subject_id) %>% 
  deframe()
no_attention

data <- data %>% filter(!exp_subject_id %in% no_attention)


participants <- participants %>%
  mutate(
    attention_check = ifelse(exp_subject_id %in% no_attention, 0, 1),
    completed = ifelse(exp_subject_id %in% incomplete, 0, 1),
    inclusion = ifelse(attention_check == 0 | completed == 0, 0, 1)
  )

participants_passed <- participants %>%
  filter(!exp_subject_id %in% incomplete,
         !exp_subject_id %in% no_attention,
         !exp_subject_id %in% no_consent,
         !exp_subject_id %in% hearing_impairement)



```


Seperate the 2 audio names from each other into different columns, and come up with a df with just audio names and similarity ratings
```{r}

similarity_sep <- data %>%
  filter(task_name %in% c("Rating", "Rating2", "Rating3", "Rating4", "Rating5", "Rating6") & audio_group != "signal") %>%
  separate_wider_delim(
    audio_name,
    delim = ".mp3",
    names = c("audio1", "audio2", "rest"),
    cols_remove = FALSE
  ) %>%
  select(!"rest") %>%
  mutate(
    audio1 = str_c(audio1,".mp3"),
    audio2 = str_c(audio2,".mp3")
  ) 


similarity_clean <- similarity_sep %>%
  select(exp_subject_id, audio1, audio2, audio_name, similarity)


```
