---
title: "What is music? Subset of stimuli presented in Exp1-3 (n = 33)"
output: html_notebook
author: Pauline Larrouy-Maestri
---

Clear and load
```{r}
rm(list = ls())

library(readxl)
library(dplyr)
library(tidyr)
library(corrplot)
library(ggplot2)
library(rstatix)
library(flextable)
library(psych)
library(ggplot2)
library(lme4)
library(lmerTest)
library(performance)
library(broom.mixed)

setwd("/Users/p.larrouy/Desktop/Music_osf/SI1")


# Reading the Excel files
df1 <- read_excel("Data/Exp1_data.xlsx",sheet = "Data_1")
df2 <- read_excel("Data/Exp2_data.xlsx",sheet = "Data_2")
df3 <- read_excel("Data/Exp3_data.xlsx",sheet = "Data_3")

# Convert the Confidence column to character type in all data frames
df1$Confidence <- as.character(df1$Confidence)
df2$Confidence <- as.character(df2$Confidence)
df3$Confidence <- as.character(df3$Confidence)

# Bind the data frames
merged_df <- bind_rows(df1, df2, df3)

# exclude participant who have NAs for the OTHER condition only :)
merged_df<-filter(merged_df, Participant !="486404") #45 
merged_df<-filter(merged_df, Participant !="487123") #97

# Check number of participants for each experiment without duplicates (here, Exp1)
unique(merged_df$Participant)
n_distinct(merged_df$Participant)
```


Select the ratings for the 33 stimuli presented in all three experiments
```{r}
stimuli_list <- c( "M_bells","M_caballo_blanco", "M_cana_verde", "M_experimental01","M_experimental04", "M_experimental05","M_experimental11", "M_experimental17", "M_faa_gani", "M_mariposa", "M_meditation", "M_miroloi", "M_piano_noise",  "M_ungava_bay", "M_zapateada_criollo", "N_bird01", "N_bird02", "N_bird03", "N_churchbells", "N_cicads", "N_coffee_grinder_new", "N_cow_bells", "N_crickets", "N_dishwasher", "N_glass_clicking", "N_guacharaca", "N_metal_percussion", "N_metal_stairs", "N_printer", "N_rhinolophe", "N_typewriter", "N_wind_chime", "N_wingflap") 

merged_df_33 <- merged_df %>%
  filter(Stim %in% stimuli_list)

unique(merged_df_33$Stim)
n_distinct(merged_df_33$Stim)

unique(merged_df_33$Participant)
n_distinct(merged_df_33$Participant)
```

*Comparison across groups: correlation matrix, anova, figure*

Aggregate per Stimuli and Condition (n = 10: "self", "other", "long", "short", "BB", "WBC", "WBR" - with test and retest for the Exp3)
```{r}
Aggdata <- aggregate(Music_answer ~ Condition + Condition_2 + Stim, data = merged_df_33, mean, na.rm = TRUE)

# Step 2: Combine Condition and Condition_2 into a single column
Aggdata <- Aggdata %>%
  unite("Combined_Condition", Condition, Condition_2, sep = "_", remove = TRUE)

unique(Aggdata$Combined_Condition)
n_distinct(Aggdata$Combined_Condition)
```

Create a wide version for correlations and repeated measures
```{r}
Cor_data <- spread(Aggdata, key = Combined_Condition, value = Music_answer)

```

Correlation between variables
```{r}
ma <- cor(Cor_data[, c("long_na","other_na","self_na","short_na","BB_retest","WBC_retest","WBR_retest","BB_test","WBC_test","WBR_test")])
ma
```
p values of these correlations
```{r}
p.ma <- cor.mtest(ma)
p.ma
```


Illustration of the correlations via correlation matrix

```{r}
ma <- cor(Cor_data[, c("self_na","other_na","short_na","long_na","BB_test","WBC_test","WBR_test","BB_retest","WBC_retest","WBR_retest")])

p.ma <- cor.mtest(ma)

# Modify column and row names
colnames(ma) <- c("self","other","short","long","BB_1","WBC_1","WBR_1","BB_2","WBC_2","WBR_2")  
rownames(ma) <- c("self","other","short","long","BB_1","WBC_1","WBR_1","BB_2","WBC_2","WBR_2")  

# Create a grayscale color palette (dark for low correlation, white for high correlation)
col_grayscale <- colorRampPalette(c("black", "white"))(200)

# Create the correlation plot with grayscale colors
corrplot(ma, 
         type = "lower", 
         col = col_grayscale,   # Apply the grayscale color palette
         tl.col = "black",      # Color of the text labels
         tl.cex = 0.9,          # Size of the text labels
         addCoef.col = "black", # Add correlation coefficients in black color
         number.cex = 0.8)      # Size of the correlation coefficient numbers


```
*Relation between bias and participants' characteristics: preprocessing, model, figure*
